---
title: "Logo"
---

Here's a quick script for creating our Warhol-esque R logo. We've adapted the offical [R logo](https://www.r-project.org/logo/), which is under a CC-BY-SA 4.0 license.

This function sets up the colour scheme. You can change `wcolors` below to anything you want.

```{r}
warhol_color <- function() {
  if (!exists("current_wcolor")) {
    current_wcolor <<- 0
  } else {
    current_wcolor <<- current_wcolor + 1
  }
  #wcolors <- c("red", "blue", "green", "yellow", "orange", "pink", "purple")
  wcolors <- c(
    "#FFF942", # yellow
    "#191464", # dark blue
    "#81A42F", # green
    "#2562D6", # royal blue
    "#CA1576", # hot pink
    "#FFF942", # yellow
    "#191464", # dark blue
    "#C91920", # red
    "#E382B3", # pink
    "#D97829", # orange
    "#572D6D", # purple
    "#76A8EA", # light blue
    "#C91920", # red
    "#18191D", # black
    "#F9F9F9", # white
    "#81A42F", # green
    "#D97829", # orange
    "#FAFEA3", # light yellow
    "#325052", # slate blue
    "#CA1576", # hot pink
    "#FAFEA3", # light yellow
    "#FFF942", # yellow
    "#C91920", # red
    "#81A42F", # green
    "#CA1576", # hot pink
    "#191464", # dark blue
    "#D97829" # orange
)
  wcolors[current_wcolor%%length(wcolors)+1]
}
```

```{r}
warhol <- function(filename = "warhol.svg", rows = 3, cols = 3) {
  logowidth <- 724
  logoheight <- 561

  svg<- paste0('<svg xmlns="http://www.w3.org/2000/svg"
               xmlns:xlink="http://www.w3.org/1999/xlink"
               preserveAspectRatio="xMidYMid"
               width="', logowidth*cols, '" height="',logoheight*rows,'"
               viewBox="0 0 ',logowidth*cols,' ',logoheight*rows,'">\n\n')

  # define the R logo in SVG
  svg <- paste0(svg, '<define>
  <rect id="bg" width="',logowidth,'" height="',logoheight,'" fill-rule="evenodd"/>
  <path id="halo" d="M361.453,485.937 C162.329,485.937 0.906,377.828 0.906,244.469
    C0.906,111.109 162.329,3.000 361.453,3.000
    C560.578,3.000 722.000,111.109 722.000,244.469
    C722.000,377.828 560.578,485.937 361.453,485.937 ZM416.641,97.406
    C265.289,97.406 142.594,171.314 142.594,262.484
    C142.594,353.654 265.289,427.562 416.641,427.562
    C567.992,427.562 679.687,377.033 679.687,262.484
    C679.687,147.971 567.992,97.406 416.641,97.406 Z"
    fill-rule="evenodd"/>
  <path id="R" d="M550.000,377.000
    C550.000,377.000 571.822,383.585 584.500,390.000
    C588.899,392.226 596.510,396.668 602.000,402.500
    C607.378,408.212 610.000,414.000 610.000,414.000
    L696.000,559.000 L557.000,559.062 L492.000,437.000
    C492.000,437.000 478.690,414.131 470.500,407.500
    C463.668,401.969 460.755,400.000 454.000,400.000
    C449.298,400.000 420.974,400.000 420.974,400.000
    L421.000,558.974 L298.000,559.026 L298.000,152.938
    L545.000,152.938 C545.000,152.938 657.500,154.967 657.500,262.000
    C657.500,369.033 550.000,377.000 550.000,377.000
    ZM496.500,241.024 L422.037,240.976 L422.000,310.026
    L496.500,310.002 C496.500,310.002 531.000,309.895 531.000,274.877
    C531.000,239.155 496.500,241.024 496.500,241.024 Z"
    fill-rule="evenodd"/>
</define>')

  current_wcolor <<- -1 # start colours at the beginning
    
  for (x in seq(0, logowidth*(cols-1), by = logowidth)) {
    for (y in seq(0, logoheight*(rows-1), by = logoheight)) {
      svg <- paste0(svg, '  <g transform="translate(',x,',',y,')">\n')
      svg <- paste0(svg, '    <use xlink:href="#bg" fill="',warhol_color(),'" />\n\n')
      svg <- paste0(svg, '    <use xlink:href="#halo" fill="',warhol_color(),'" />\n\n')
      svg <- paste0(svg, '    <use xlink:href="#R" fill="',warhol_color(),'" />\n\n')
      svg <- paste0(svg, '  </g>\n\n')
    }
  }

  svg <- paste0(svg, '</svg>')

  write(svg, file = filename)
}
```

Just run the following code to make a new logo:

```{r, eval = F}
warhol("images/logo_square.svg", 3,3)
```

![You can make a square version](images/logo_square.svg)

```{r, eval = F}
warhol("images/logo_wide.svg", 1,6)
```

![Or a wide version](images/logo_wide.svg)


```{r, eval = F}
warhol("images/logo_too_many.svg", 30,30)
```

![Or even a version with `r 30*30` Rs](images/logo_too_many.svg)

